---
title: "Assignment 1 Macro"
author: "Mohamed Sabath_COHNDDS23.1F-007"
date: "2023-09-08"
output: html_document
---

```{r}
setwd("D:/Data science NIBM/HND/Macroeconomics/Assignment 1")
```


step 1 :Data Exploration 

```{r}
# Load the readxl package
library(readxl)

# Load the data from the Excel file
data <- read_excel("CA 1 Data set.xlsx")


# first few rows of the dataset
head(data)


```


Step 2: Data Loading and Cleaning 

```{r}
# Checking for missing values
colSums(is.na(data))

```
it has no null values.


Step 3 : Descriptive Statistics and Visualization.
```{r}
# Getting basic information about the dataset
str(data)

# Getting summary statistics for numerical columns
summary(data)

# Getting the list of column names
colnames(data)

```
```{r}
library(ggplot2)
ggplot(data, aes(x = Year, y = EI,)) +
  geom_line(color = "darkgreen") +
  scale_color_brewer(palette = "Set1") +
  labs(title = "Line Graph for Exports and import Ratio",
       x = "Year",
       y = "Exports and Import Ratio") +
  theme_minimal()

```
#Here it looks like by the year 2012 it has reached the ratio of 1.0 which means imports = exports.

```{r}
# Customized line graph with points
ggplot(data, aes(x = Year, y = `GDP ($Mn)`)) +
  geom_line(color = "#b042f5",size = 1) +
  geom_point(shape = 16, size = 1, color = "red") +
  labs(x = "Year", y = "GDP ($Mn)", title = "Line Graph with Points for the GDP")
  scale_x_continuous(breaks = seq(min(data$Year), max(data$Year), by = 2))

```
Seems like by 1984 the gdp was 300mn. But after that we can see a continous decreament on the gdp until 1885.


```{r}
ggplot(data, aes(x = IH,)) +
  geom_histogram(fill = "pink", colour = "black") +
  labs(title = "Histogram of Per Capita income per worker hour")
```
```{r}
#  line graph with points
ggplot(data, aes(x = Year, y = `IH`)) +
  geom_line(color = "#f5bf42",size = 1) +
  geom_point(shape = 16, size = 1, color = "red") +
  labs(x = "Year", y = "Per Capita income per work hour", title = "Line Graph with Points for the percapita income per work hour")
  scale_x_continuous(breaks = seq(min(data$Year), max(data$Year), by = 2))
```
```{r}
library(ggplot2)
library(reshape2)

cor_matrix <- cor(data[, c("ST", "EI", "IH", "GDP ($Mn)","BI ($Mn)")])

cor_matrix_melted <- reshape2::melt(cor_matrix)

ggplot(data = cor_matrix_melted, aes(x = Var1, y = Var2)) +
  geom_tile(aes(fill = value), color = "white") +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  scale_fill_gradient(low = "#42f5d7", high = "#9b42f5") +
  labs(title = "Correlation Heatmap with Coefficient Values",
       x = "Variable 1",
       y = "Variable 2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
library(scatterplot3d)
# Create a 3D scatter plot
scatterplot3d(data$ST, data$EI, data$`GDP ($Mn)`,
  color = "blue", pch = 19, type = "h", main = "3D Scatter Plot",
  xlab = "Subsidies per capita", ylab = "Exports and Imports", zlab = "GDP"
)
```

```{r}
# Define custom colors
colors <- c('blue', 'green', 'red', 'cyan')

# Setting up a multi-panel layout
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))

# Create time series plots for key variables with custom colors
plot(data$Year, data$ST, type = 'o', col = colors[1], xlab = 'Year', ylab = 'ST (Subsidies per Taxes Collected)', main = 'Time Series Plot of ST')
plot(data$Year, data$EI, type = 'o', col = colors[2], xlab = 'Year', ylab = 'EI (Exports to Imports)', main = 'Time Series Plot of EI')
plot(data$Year, data$IH, type = 'o', col = colors[3], xlab = 'Year', ylab = 'IH (Per Capita Income per Worker Hour)', main = 'Time Series Plot of IH')
plot(data$Year, data$`GDP ($Mn)`, type = 'o', col = colors[1], xlab = 'Year', ylab = 'Value ($Mn)', main = 'Time Series Plot of GDP')
lines(data$Year, data$`BI ($Mn)`, type = 'o', col = colors[2])
legend('topright', legend = c('GDP ($Mn)', 'BI ($Mn)'), col = colors[1:2], lty = 1)

# Add overall title
mtext('Time Series Plots for all the Key Variables', outer = TRUE, cex = 1.2)

# Reset the plotting parameters
par(mfrow = c(1, 1), oma = c(0, 0, 0, 0))

```



```{r}
# Load the ggplot2 package
library(ggplot2)

# Create a density plot for Business Investments
ggplot(data, aes(x = data$`BI ($Mn)`)) +
  geom_density(fill = "lightblue", alpha = 0.7) +
  labs(x = "Business Investments ($Mn)", y = "Density", title = "Density Plot for Business Investments")


```

The distribution shape seems not like bell shaped (symmetric). 

```{r}
ggplot(data, aes(x = `GDP ($Mn)`, y = `BI ($Mn)`)) +
  geom_point(color = "blue", shape = 3) +  
  labs(title = "Scatterplot of GDP and Business investment variable")
```


#The scatterplot above shows the relationship/ correlation between GDP and Business investments. From the chart, a positive linear correlation is visible (When one variable increases, the other tend to increase as well). Inother word we can say that the both variable has a good positive relationship.





Step 4: Analysis and Interpretation.
          1.Hypothesis testing.

hypothesis = is there any relationship between Export and index and the Bisiness Investment.

Null hypothesis(H0) = There is no relationship between Export and index and the Bisiness Investment.
Alternate Hypothesis (H1) = There is a relationship between Export and index and the Bisiness Investment.



```{r}
lin_mod_EI_BI <- lm(EI~`BI ($Mn)`, data = data)
summary(lin_mod_EI_BI)
```
So here we can see that the p value is less that 0.05. So that means we can't reject the null hypothesis. So we have to go with the alternative hypothesis. which means There is a relationship between Export and index and the Business Investment.



        

